---
title: "Lecture 1"
author: "Jung-Jin Lee"
date: "Jan 7, 2019"
output:
  xaringan::moon_reader:
    css: ["default", "default-fonts", "MLAS.css"] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      
---
```{r, message = F, echo = F, warning = F}
library(tidyverse)
library(knitr) 
opts_chunk$set(
  tidy = FALSE,
  cache = FALSE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  dpi = 300,
  fig.align = "center",
  fig.width = 3,
  fig.height = 3
  )
```

# Installing R/RStudio

- To install R, go to https://www.r-project.org/ and follow instruction there
    - Click on CRAN under Download on the left
    - Choose the closest mirror site (e.g. http://lib.stat.cmu.edu/R/CRAN/)
    - Download R for your platform (Windows/Mac/Linux)
    
- To install RStudio (an integrated development environment for R), go to https://www.rstudio.com/ and follow instruction

---

# R as a calculator

```{r}
5 + 3
```

```{r}
# everything after # (pound sign) is ignored
# useful for comments
2 * 5 # sum of 2 and 5
```
 
---

# Numerical variables

Variables can be used to store numerical values. 

- A variable can be a combination of letters (case sensitive), digits, period (.) and underscore (_).
- It must start with a letter or a period. If it starts with a period, it cannot be followed by a digit.
- Reserved words in R, such as `TRUE`, cannot be used as variables.

```{r}
Num1 <- 5; Num2 <- 3 # semicolon (;) can be used to separate lines
sum_num <- Num1 + Num2
prod.num <- Num1 * Num2
print(sum_num) 
print(prod.num)
```

---

# Next

```{r, out.width = "50%", out.height = "20%"}
g <- ggplot() + 
  geom_point(aes(x = 1:4, y = 2:5)) +
  theme(aspect.ratio = 1)
print(g)
```


