---
title: "Lecture 1"
author: "Jung-Jin Lee"
date: "Jan 7, 2019"
output:
  xaringan::moon_reader:
    css: ["default", "default-fonts", "MLAS.css"] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      
---
```{r, message = F, echo = F, warning = F}
library(tidyverse) 
library(knitr) 
opts_chunk$set(
  tidy = FALSE,
  cache = FALSE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  dpi = 300,
  fig.align = "center",
  fig.width = 3,
  fig.height = 3
  )
```

## Installing R/RStudio

- To install R, go to https://www.r-project.org/ and follow instruction there
    - Click on CRAN under Download on the left
    - Choose the closest mirror site (e.g. http://lib.stat.cmu.edu/R/CRAN/)
    - Download R for your platform (Windows/Mac/Linux)
    
- To install RStudio (an integrated development environment for R), go to https://www.rstudio.com/ and follow instruction

---

## R as a calculator

```{r}
5 + 3
```

```{r}
# everything after # (pound sign) is ignored
# useful for comments
2 * 5 # product of 2 and 5
```

```{r}
3^2
```

```{r}
3**2
```

---

## Variables

Variables can be used to store numerical values. 

- A variable can be a combination of letters (case sensitive), digits, period (.) and underscore (_).
- It must start with a letter or a period. If it starts with a period, it cannot be followed by a digit.
- Reserved words in R, such as `TRUE`, cannot be used as variables.

```{r}
Num1 <- 5; Num2 <- 3 # semicolon (;) can be used to separate lines
sum_num <- Num1 + Num2
prod.num <- Num1 * Num2
print(sum_num) 
print(prod.num)
```

---

## Characters

Variables can also be used to store character values.

```{r}
first_name <- "John" # use quotation marks for character strings
last_name <- 'Smith' # single quotation also works
print(first_name)
print(last_name)
```

---

## Functions

One or more values can produce new one using R functions.

```{r}
a <- 1; b <- 3
s <- sum(a, b) # sum() is a function that computes the sum
print(s)
```

```{r}
first_name <- "John" 
last_name <- "Smith"
# paste() puts together multiple character strings
full_name <- paste(first_name, last_name) 
print(full_name)
```

For detailed usage of a function, type `help(function_name)`, e.g. `help(paste)` or `?function_name`, e.g. `?paste`.

---

## Vectors

Multiple numerical values or characters can be saved as a single variable using an R function `c()`.

```{r}
vec1 <- c(2,3,5,7,1,4)
print(vec1)
```

```{r}
vec2 <- c("Drexel", "Univ", "School", "of", "Medicine")
print(vec2)
```

---

## Functions applied to a vector

```{r}
length(vec1)
```

```{r}
sum(vec1)
```

```{r}
DUSoM <- paste(vec2, collapse = " ")
print(DUSoM)
```

```{r}
DUSoM2 <- paste(vec2, collapse = "_")
print(DUSoM2)
```

---

## Vector manipulation

Use brackets `[]` to extract components of a vector.

```{r}
a <- vec1[3]
print(a)
```

```{r}
dm <- vec2[c(1, 5)]
print(dm)
```

---

## Special vectors -- part I

Exercise: compute the sum $1 + 2 + 3 + \cdots + 10$.

```{r}
w1 <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
sum(w1)
```

```{r}
# colon (:) can be used to generate a vector consisting of 
# consecutive numbers
w2 <- 1:10
print(w2)
sum(w2)
```

---

## Special vectors -- part II

Exercise: compute the sum $2 + 4 + 6 +\cdots + 20$.

```{r}
# one can use a function seq()
# type ?seq to get help
w3 <- seq(2, 20, by = 2)
sum(w3)
```

Alternatively, one can use the previous vector:

```{r}
w4 <- 2*w1 # product of a single numner and a vector
print(w4)
sum(w4)
```


---

## Next

```{r, out.width = "50%", out.height = "20%"}
g <- ggplot() + 
  geom_point(aes(x = 1:4, y = 2:5)) +
  theme(aspect.ratio = 1)
print(g)
```


